Cover letter for the IJRR submission \\ ``An efficient acyclic contact planner for multiped robots''

### 2016-01-22 body/overview.tex
3.1

@@

\medskip
\textbf{A robot} is a kinematic chain $R$, comprising \mbox{$n + r$} degrees of freedom (DOFs), with $r \geq 6$. $r$ denotes the world coordinates of the root of the robot,
as well as extra DOFs not related to the limbs (for instance, HRP-2 has two extra DOFS in the torso, such that we have $r=8$). Abusively, for simplicity,
we refer to the root as the kinematic chain described by all these DOFS.
$R$ is composed of $l$ limbs $R^k, 1 \leq k \leq l$, attached to the root.
It is described by a configuration $\mathbf{q} \in  \mathbb{R}^r \times \mathbb{R}^n$.
We define some relevant projections of $\mathbf{q}$:
\begin{itemize}
	\item $\mathbf{q}^k$ denotes the configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ denotes the vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ denotes the world coordinates of the root of the robot $R$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing the root of the robot is denoted $W^0$.
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 The reachable workspace of a limb $R^k$ is denoted $W^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\},
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0 = \mathbf{0}$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define $W = \bigcup_{k=1}^{l}W^k$, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ translated and rotated according to the root configuration $\mathbf{q}^{0}$.

>>>>

In the remainder of this paper, we use the terms \gls{contact feasible} and \gls{equilibrium feasible} to qualify either a root configuration, a whole-body configuration, or a set
of such configurations.

@@

\medskip
\textbf{A robot} is a kinematic chain $R$, that we divide into different parts: a root $R^0$, and $l$ limbs $R^k, 1 \leq k \leq l$, attached to the root.
The root has $r \geq 6$ DOFS: for instance, HRP-2 has two extra DOFS in the torso, such that we have $r=8$.
Therefore $R$ is fully described by a configuration $\mathbf{q} \in  \mathbb{R}^{r+n}$.
We define some relevant projections of $\mathbf{q}$:
\begin{itemize}
	\item $\mathbf{q}^k$ denotes the configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ denotes the vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ denotes the world coordinates of the root $R^0$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing $R^0$ is denoted $W^0$.
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 The reachable workspace of a limb $R^k$ is denoted $W^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\}
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0 = \mathbf{0}$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define \mbox{$W = \bigcup_{k=1}^{l}W^k$}, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ for the root configuration $\mathbf{q}^{0}$.

### 2016-01-22 body/rbprm.tex
typos
3.1

% !TEX root =  ../main.tex
\section{Computing a guide path in $C_{reach}$ ($\mathcal{P}_1$) }
\label{rbprm}

We consider the issue of computing a normalized guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow \mathbb{R}^r$ for the root of a multiped robot, connecting user-defined start and goal configurations. As said in the previous section, the goal is to enforce that the guide can be discretized into a sequence of \gls{contact feasible} configurations. We denote by \gls{ccontact} the contact sub-manifold of the robot.

We say that a root configuration $\mathbf{q}^{0}$ is \gls{contact feasible} if it belongs to the set:
\begin{equation}
      C_{contact}^0= \{ \mathbf{q}^{0}: \exists \mathbf{q}^{\overline{0}}, \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact} \}
    \label{eq:pi}
\end{equation}

%~ The set of all \textit{geometrically feasible} root configurations is denoted by $C_{reach}$.
By extension, a path $\mathbf{q}^0(t)$ is \gls{contact feasible} if $\forall t \in [0,1], \mathbf{q}^0(t) \in C_{contact}^0$.

For this stage, and for this stage only, we make the assumption that for \gls{cluttered} problems, a \gls{contact feasible} path can be transformed into a \gls{equilibrium feasible} path.
Under this assumption, for a two-stage acyclic contact planner to be exact and complete, we need the combination of two conditions on a guide path generator: all the generated paths must be \gls{contact feasible} (sufficient condition); the guide planner must be complete, i.e. it must be able to discover any \textit{contact feasible} path (necessary condition).

In this Section we characterize \gls{ccontact0} theoretically with a necessary and a sufficient condition for belonging to the set.
We then approximate \gls{ccontact0} as a space $C_{reach}$, where we can efficiently sample configurations.
We finally describe our implementation of an efficient RRT planner in $C_{reach}$.

@@

Therefore we can define an optimistic approximation  $C^0_{Nec} \supset$ \gls{ccontact0} defined as: 
\begin{equation}
C_{Nec}^0 = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \wedge & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
The inclusion \gls{ccontact0} $\subset C^0_{Nec}$ is straightforward, but it is very important.
Planning in $C^0_{Nec}$ allows a strong reduction of the search space, while not discarding 
any root configuration that could lead to a solution.

@@

However, between $W^0$ and $B^{Suf}$, a trade-off can be found between a necessary and a sufficient condition. We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $C_{Nec}$. We thus define $s \in \mathbb{R}^+$, since values inferior to $1$ can only introduce more errors.
The higher $s$, the closer this \gls{reachc} is to being sufficient, and if $s=1$, the search is complete (in the sense that all \gls{contact feasible} configurations can be found). The parametrization of $s$ defines a trade-off between these two desirable properties. $s$ can be chosen by hand, or automatically as explained in Section~\ref{sec:params}.
In the rest of the paper, we consider that an appropriate value $s^*$ of $s$ is chosen, and define \gls{creach0} $= C^0_{s^*}$.

>>>>

% !TEX root =  ../main.tex
\section{Computing a guide path in $C^0_{reach}$ ($\mathcal{P}_1$) }
\label{rbprm}

We consider the issue of computing an \gls{equilibrium feasible} guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow$ \gls{ccequil0} for the root of a multiped robot, connecting user-defined start and goal configurations.

Again, we assume that, for \gls{cluttered} problems, most of the times \contactfeasibility\, implies \equilibriumfeasibility.
Under this assumption, our goal is to find a \gls{contact feasible} guide path.

To generate such a path efficiently, ideally we need to only sample \gls{contact feasible} root configurations.
This requires exhibiting a necessary and sufficient condition for \contactfeasibility\,that does not need to compute the whole-body configuration.
Since we do not know such a condition, we may turn our attention towards either an only-necessary or an only-sufficient condition.

Only-necessary conditions are appealing because they preserve the completeness of the search, while reducing 
the search space: they provide an outer-approximation of \gls{ccontact0}.
On the other hand, only-sufficient conditions provide the guarantee that any configuration that satisfies them is indeed \gls{contact feasible}:
they provide an inner-approximation of \gls{ccontact0}.

In practice, the only-necessary and only-sufficient conditions that we can provide are trivial and provide rather inaccurate approximations of \gls{ccontact0}.
Therefore, we propose a compromise between them: the \textit{reachability condition}, which is computationally efficient
and it provides a rather accurate approximation of \gls{ccontact0}. % for \gls{cluttered} problems.

%~ As said in the previous section, the goal is to enforce that the guide can be discretized into a sequence of \gls{equilibrium feasible} configurations. We denote by \gls{ccequil} the equilibrium sub-manifold of the robot.

%~ We say that a root configuration $\mathbf{q}^{0}$ is \gls{equilibrium feasible} if it belongs to the set:
%~ \begin{equation}
      %~ C_{Equil}^0= \{ \mathbf{q}^{0}: \exists \mathbf{q}^{\overline{0}}, \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{Equil} \}
    %~ \label{eq:pi}
%~ \end{equation}

%~ The set of all \textit{geometrically feasible} root configurations is denoted by $C_{reach}$.
%~ By extension, a path $\mathbf{q}^0(t)$ is \gls{equilibrium feasible} if $\forall t \in [0,1], \mathbf{q}^0(t) \in C_{contact}^0$.

%~ In this Section we characterize \gls{ccontact0} theoretically with a necessary and a sufficient condition for belonging to the set.
%~ We then approximate \gls{ccontact0} as a space $C_{reach}$, where we can efficiently sample configurations.
%~ We finally describe our implementation of an efficient RRT planner in $C_{reach}$.

@@

Therefore we can define an optimistic approximation  $C^0_{Nec} \supset$ \gls{ccontact0} defined as: 
\begin{equation}
C_{Nec}^0 = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
The inclusion \gls{ccontact0} $\subset C^0_{Nec}$ is straightforward, but it is very important.
Planning in $C^0_{Nec}$ allows a strong reduction of the search space, while not discarding 
any root configuration that could lead to a solution.

@@

However, between $W^0$ and $B^{Suf}$, a trade-off can be found between a necessary and a sufficient condition. We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $C_{Nec}$. We thus define $s \in \mathbb{R}^+$, since values inferior to $1$ can only introduce more errors.
The higher $s$, the closer this \gls{reachc} is to being sufficient, and if $s=1$, the search is complete (in the sense that all \gls{contact feasible} configurations can be found). The parametrization of $s$ defines a trade-off between these two desirable properties. $s$ can be chosen by hand, or automatically as explained in Section~\ref{sec:params}.
In the rest of the paper, we consider that an appropriate value $s^*$ of $s$ is chosen, and define \gls{creach0} $= C^0_{s^*}$.
